---
title: "US Polling Places 2012-2020"
subtitle: "Tracking Changes in US Polling Places from 2012 to 2020"
author: 
  - name: "DataDuo"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "This project aims to analyze the changes in US polling places over the period from 2012 to 2020, examining trends and shifts in polling location availability at the county level."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(lubridate)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
polling_places <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-01-16/polling_places.csv')
glimpse(polling_places)

dim(polling_places)
names(polling_places)
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

\# Dataset Description

The dataset contains information on US polling places from 2012 to 2020, including locations, years active, demographic information of precincts, voter turnout, and changes in polling place locations over time. The dataset has rows and columns. It was chosen to analyze trends and factors influencing polling place changes and voter accessibility

## Questions

The two questions you want to answer.

Question 1 : What are the trends in the number of polling places over time, both nationally and across different states from 2012 to 2020?

Question 2 :

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
